GPU-Beschleunigung der Material Point Method

### 2019-01-18 mscthesis.tex
performance tables

Parallel reduction solves assuming a \texttt{unary\_op(x)} directly after memory load:
\begin{itemize}
  \item CFL-condition of chapter \ref{sec:cfl} with operators
  \begin{itemize}
  \item \texttt{unary\_op(x) := length2(x)} and
  \item \texttt{binary\_op(x,y) := max(x,y)}.
  \end{itemize}
\item Discretized conservation of governing equations between time steps or transfers in chapter \ref{sec:mass} and \ref{sec:apic} with
\begin{itemize}
  \item \texttt{unary\_op(x) := x} and
  \item \texttt{binary\_op(x,y) := x + y}.
  \end{itemize}
\item Calculate the maximum count of particles of all grid nodes within a block (\textit{per-block parallel reduction} \ref{sec:active_blocks}):
\begin{itemize}
  \item \texttt{unary\_op(x) := x} and
  \item \texttt{binary\_op(x,y) := max(x,y)}.
\end{itemize}
\end{itemize}

@@

Starting a dispatch for every block only the active blocks will now write their index at the scan's position. This process is not coalesced. Coalescing will still be guarenteed for all nodes within the block for the P2G-Transfers and G2P-Transfer. The only non-coalesced access within these shaders is the maximum per block count, which there are really few of in comparison.

\section{Evaluation}\label{sec:eval}
\subsection{Verifying Results}
\begin{itemize}
  \item Conservation equations
    \cite{MPM:APIC}
  \item Singular Value Decomposition
\end{itemize}

\textbf{SoA vs. AoS}:

\begin{table}
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Description &  $\Delta t_c (\mu s)$ & Speedup &VRAM     & SM     & L2     & Tex \\\hline
    AoS         & 278                   & -       &77.7\%   & 7.3\%  & 30.3\% & 14.7\% \\\hline
    SoA         & 122                   & 2.26x   &75.4\%   & 14.3\% & 29.4\% & 17.9\% \\
    \hline
  \end{tabular}
  \caption{Map operation with one instruction of form \texttt{float x = length2( (vec4) y)} on $1024 \times 1024$ \texttt{vec4}.}\label{table:map_1}
\end{table}

\begin{table}
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Description &  $\Delta t_c (\mu s)$ & Speedup &VRAM   & SM     & L2     & Tex \\\hline
    AoS         & 292                   & -       &61.3\% & 11.4\% & 41.8\% & 20.4\% \\\hline
    SoA         & 252                   & 1.16x   &75.4\% & 14.3\% & 29.4\% & 17.9\% \\
    \hline
  \end{tabular}
  \begin{tabular}{ | l | l | l | l | }    \hline
    Description &  L2-Hit & Read Active \\\hline
    AoS   & 53.8\%        & 48.9\%      \\\hline
    SoA   & 20.0\%          & 62.3\%      \\
    \hline
  \end{tabular}
  \caption{Map operation with two instructions of form \texttt{float x = length2( (vec4) y)} on $1024 \times 1024 \times 2$ \texttt{vec4}.}\label{table:map_2}
\end{table}

>>>>

Parallel reduction solves assuming a \texttt{unary\_op(x)} directly after memory load:
\begin{itemize}
  \item CFL-condition of chapter \ref{sec:cfl} with operators
  \begin{itemize}
  \item \texttt{unary\_op(x) := length(x)} and
  \item \texttt{binary\_op(x,y) := max(x,y)}.
  \end{itemize}
\item Discretized conservation of governing equations between time steps or transfers in chapter \ref{sec:mass} and \ref{sec:apic} with
\begin{itemize}
  \item \texttt{unary\_op(x) := x} and
  \item \texttt{binary\_op(x,y) := x + y}.
  \end{itemize}
\item Calculate the maximum count of particles of all grid nodes within a block (\textit{per-block parallel reduction} \ref{sec:active_blocks}):
\begin{itemize}
  \item \texttt{unary\_op(x) := x} and
  \item \texttt{binary\_op(x,y) := max(x,y)}.
\end{itemize}
\end{itemize}

@@

Starting a dispatch for every block only the active blocks will now write their index at the scan's position. This process is not coalesced. Coalescing will still be guarenteed for all nodes within the block for the P2G-Transfers and G2P-Transfer. The only non-coalesced access within these shaders is loading the maximum per block count from VRAM, which there are really few of in comparison.

\section{Evaluation}\label{sec:eval}
\subsection{Verifying Results}
\textbf{Parallel reduction}, \textbf{map} and \textbf{scan} are tested against \texttt{std::transform\_reduce} as it supplies the same functionality as a fully reduced parallel reduction. It is standard since C++17.
\begin{table}[h]
  \begin{tabular}{|l|r|r|r|}\hline
    Shader       & \texttt{unary\_op} & Total abs. error & Total rel. error \\\hline
    Map (float)   & x          & 0.0    & 0.0      \\\hline
    Map (vec4,dvec4)   & length(x)  &0.0625 & 7.94e-08 \\\hline\hline
    MapReduce (float)& x  &0.0183 & 3.51e-08 \\\hline
    MapReduce (double)& x  &1.78e-08 & 3.42e-14 \\\hline
    MapReduce (vec4)& length(x)  &212.813 & 2.71e-04 \\\hline
    MapReduce (dvec4)& length(x)  &1.58e-05 & 2.02e-11 \\\hline
  \end{tabular}
  \caption{Error of various shaders on 1024*1024 random elements. Scalars are just between [0.0-1.0]. The vectors have length between [0.0-1.0].}
\end{table}

\begin{itemize}
  \item Conservation equations
    \cite{MPM:APIC}
  \item Singular Value Decomposition
\end{itemize}
 \subsubsection{Performance}
\textbf{SoA vs. AoS}:
dvec4 same
\begin{table}
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Layout &  $\Delta t_c (\mu s)$ & Speedup &VRAM     & SM     & L2     & Tex \\\hline
    AoS         & 278                   & -       &77.7\%   & 7.3\%  & 30.3\% & 14.7\% \\\hline
    SoA         & 122                   & 2.26x   &75.4\%   & 14.3\% & 29.4\% & 17.9\% \\
    \hline
  \end{tabular}
\caption{Timings betweens different buffer layouts. Map operation with one instruction \texttt{float x = length(vel)} on $1024 \times 1024$ \texttt{vec4} buffer storing position and velocity (\texttt{vel})}.}\label{table:map_1}
\end{table}

\begin{table}
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Layout &  $\Delta t_c (\mu s)$ & Speedup       &VRAM             & L2              & L2-Hit & Read Active \\\hline
    AoS         & 292                   & -        &61.3\%           & \textbf{41.8}\% & 53.8\% & 48.9\% \\\hline
    SoA         & 252                   & 1.16x    &\textbf{75.4}\%  & 29.4\%          & 20.0\% & 62.3\% \\
    \hline
  \end{tabular}
\caption{Timings betweens different buffer layouts. Map operation with instructions \texttt{float x = length(vel), float y = length(pos)} on $1024 \times 1024$ \texttt{vec4} buffer storing position(\texttt{pos}) and velocity (\texttt{vel})}.}\label{table:map_1}
\end{table}

\begin{enumerate}
Occupancy!
\end{enumerate}
\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Method         &  $\Delta t_c (\mu s)$ & Speedup &VRAM              & SM              & Top Warp-Stall Reas.\\\hline
    Interleaved    & 305                   & -       &23.0\%            & \textbf{60.9}\% & Short Scoreb.(17.2\%)\\\hline
    Sequential     & 141                   & 2.16x   &\textbf{49.8}\%   & 37.1\%          & Long Scoreb.(17.3\%) \\\hline
    Seq. (2x)      & 100                   & 3.05x   &69.5\%            & 26.2\%          & Long Scoreb.(80.1\%)\\\hline
    Seq. (128x)    & 98                    & 3.1x    &\textbf{72.9}\%   & 16.9\%          & Long Scoreb.(84.4\%)\\\hline
    Seq. (256x)    & 101                   & 3.0x    &66.4\%            & 14.6\%          & Long Scoreb.(76.9\%)\\
    \hline
  \end{tabular}
  \caption{Optimization of one parallel reduction dispatch on $1024\times1024$ instruction \texttt{float x= length(vec4 y)}. Note that the interleaved addressing intentionally does bank conflicts to show performance impacts. The number in parentheses states the number of sequential elements each thread loads.}\label{table:map_reduce}
\end{table}

\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r | r | r |}    \hline
    Method              &  $\Delta t_c$         & Speedup &VRAM              & SM      & L2     & Top Stall Reas.\\\hline
    Conflicts           & 748                   & -       &17.5\%            & 51.7\%  & 8.2\%  & Barrier(49.5\%)\\\hline
    No confl.           & 571                   & 1.31x   &25.7\%            & 50.1\%  & 5.4\%  & Barrier(31.9\%)\\\hline
    part. confl.        & 546                   & 1.36x   &23.5\%            & 50.8\%  & 6.0\%  & Barrier(30.5\%) \\\hline\hline
    Seq. (2x)           & 323                   & 2.31x   &41.7\%            & 43.0\%  & 17.2\% & S. Scoreb.(32.0\%)\\\hline
    unrolled            & 258                   & 2.90x   &53.8\%            & 42.2\%  & 22.0\% & S. Scoreb.(32.0\%)\\\hline\hline
    Seq. (4x)           & 311                   & 2.40x   &56.8\%            & 22.7\%  & 37.9\% & L. Scoreb.(43.3\%)\\\hline
    unrolled            & 297                   & 2.52x   &61.1\%            & 21.2\%  & 40.6\% & L. Scoreb.(49.9\%)\\
    \hline
  \end{tabular}
  \caption{Optimization of one scan dispatch on $4 \times 1024\times1024$ integers with \texttt{binary\_op(x,y) := x+y}. $\Delta t_c$ in $\mu s$. }\label{table:scan}
\end{table}

\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |r|}    \hline
    Ordering            &  $\Delta t_c (\mu s)$ & Speedup &VRAM              & SM      & L2     & L2-Hit \\\hline
    Random              &  1,516                & -       &25.0\%            & 3.4\%   & 9.1\%  & 10.8\% \\\hline
    Deep sort           &    218                & 6.95x   &75.3\%            & 24.4\%  & 35.0\% & 37.8\% \\
    \hline
 \end{tabular}
\caption{Order dependency of binning of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid.}\label{table:bin}
\end{table}

\begin{table}
  \begin{tabular}{ | l | r | r | r | r | r |r|}    \hline
    Best Case(*)          &  $\Delta t_c$         &VRAM  & Top Stall Reas. & Read & Write\\\hline
    Index                 &  1,516                &25.6\%& L. Scoreb.(78.7\%) &&\\\hline
    Full(8 vec4)     &  1,731                &75.3\%& L. Scoreb.(92.2\%) &&\\
    \hline
    \hline
    Worst Case(*)          &  $\Delta t_c$         &VRAM              & Top Stall Reas.    &Read & Write\\\hline
    Index                  &  1,516                &26.0\%            & L. Scoreb.(73.5\%) & &\\\hline
    Deep Sort(8 vec4)      &  9,847                &24.0\%            & L. Scoreb.(82.2\%) & &\\
    \hline
 \end{tabular}
\caption{Order dependency of binning of $1024\times1024$ randomly positioned particles in a $128\times128\times128$ grid.}\label{table:bin}
\end{table}
